### 2017-05-11 paper.tex
updated references and related work

\documentclass[11pt]{sensys-proc}

\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\graphicspath{{./figures/} {./plots/}}

\usepackage{balance}
\usepackage{comment}
\usepackage{listings}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{wrapfig}
\newcommand{\chain}{Chain\xspace}
\numberofauthors{3}

%\author{
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
%% e-mail address with \email.
%\alignauthor Alice Security \\
%        \affaddr{Department of Computer Science}\\
%        \affaddr{University of Southern California}\\
%       \email{alice@example.edu}
%\alignauthor Bob Privacy \\
%    \affaddr{Networked Embedded Systems Group}\\
%    \affaddr{Swedish Institute of Computer Science}\\
%    \email{bob@example.se}
%}

\title{TAPIR: Threaded Application Programming on Intermittent Resources}

\begin{document}

\maketitle

%Only including an abstract since it looks like the submission site wants one...
\begin{abstract}
Killer abstract.
\end{abstract}

\section{Introduction/Background} -Amanda
  \label{sec:intro}
  Chain, DINO\\
  Multithreading\\
  Intermittance Review\\
  RMW\\
  Example Citation\cite{Chain, Mementos, Aware, Dino}!

%Sample of adding a picture across the top of a split page...
\begin{figure*}
\centering
\begin{minipage}[b]{0.49\textwidth}
  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
\caption{Capybara power system overview}\label{label-a}
\end{minipage}\hfill
\begin{minipage}[b]{0.49\textwidth}
  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
\caption{Effect of different power modes}\label{label-b}
\end{minipage}
\end{figure*}

More text after the picture to show how that looks...

\section{Related Work} -Emily
In addition to the division of a program into tasks, as described above, prior work
explores several different strategies for preserving state across reboots. The approaches
range from fully independent of the programmer~\cite{ratchet, dewdrop} at one extreme, to programmer
defined task boundaries~\cite{dino} at the other.  Compiler driven approaches such as the
method used in ~\cite{Mementos} allow the programmer to specify high level behavior of the
checkpointing system. A key similarity among the prior work is it
does not lend itself well to multi-programming let
alone multi-threading. In fact, many of these works state the assumption that the device
hardware cannot support a conventional OS, and thus focus on single threaded programs. 
Despite the availability of (relatively) inexpensive nonvolatile memory technology such as
FRAM~\cite{quickrecall}, the the energy and storage overhead of checkpointing multiple
thread states would limit the maximum per-thread checkpoint size.

Furthermore, if a multiprogrammed application were implemented using previous
checkpointing strategies, reasoning about thread interleavings becomes  difficult to
reason about because of unpredictable power failures. Any conventional timeout based
thread switching approach will be challenging in an intermittent system, but systems that
allows dynamic checkpointing~\cite{hibernus}, or dynamic task formation~\cite{Dino} become
even more difficult to program. Particularly in systems like ~\cite{Dino} where the
programmer is responsible for placing task boundaries, there is a risk that programmers
will either overuse task boundaries in an effort to preserve any given thread's state, or
underprovision task boundaries such that progress is never made because of context
switching that dynamically expands the number of instructions between task boundaries.

While prior work has stopped short of suggesting an operating system for intermittently
powered devices, a number of different works touch on the need for these devices to
perform multiple tasks in the same workload. Buettner, Greenstein and Wetherall motivate
their runtime, Dewdrop, based on difference in energy requirements for the tasks that a
CRFID needs to perform~\cite{dewdrop}. The energy aware memory mapping in ~\cite{Aware}
discusses the challenge of handling nondeterministic interrupts received by intermittent
sensing platforms when a peripheral must be serviced immediately and control flow deviates
from single threaded execution. From a hardware standpoint, the federated energy scheme in
~\cite{ufop} provides a platform for removing the energy dependence between an MCU and its
peripherals. Federated energy also simplifies a programmer's reasoning about reading from
a single peripheral. However, federating energy does not help a programmer describe how to
handle {\em multiple} peripherals. Consider a listing in
~\cite{ufop} reporoduced in Figure ~\ref{copy}, and imagine the level of understanding
of the hardware behavior a programmer would have to employ to, say, read one sensor if it
is available but read from and proceed to process data from a different sensor if it is not.

Operating system work\\
Want to do multiple things- i.e. federated energy \\
Dewdrop- energy aware task scheduling\\
TU delft guys...\\
Energy Align\\
TinyOS stuff in wireless sensor networks\\

\section{Chain Discussion} -NEIL

Arrays in channels and why that's bad\\
MAX\_NUM\_DIRTY\_FIELDS\\
Exploiting the fact that there isn't a compiler\\

\section{What we did} % TODO change title
Scheduler\\
Mutexes - Emily\\
User programming model for threads - Neil\\
Round Robin (alternatives, on reboot)\\
SIMD vs MIMD\\
SIMD w/ channels (you don't get your registers...etc -
    just channels, since intermittent\\
\subsection{Interrupts}
\subsubsection{Motivation}
Ideally, multithreading would extend to a programming model that can utilize
interrupts in a producer/consumer fashion (i.e. a thread is waiting on data
from an external source, the external source fires an interrupt, the thread
samples the external source and stores the received data in a buffer, and
the thread returns to its original task). However, the inherent
non-deterministic
nature of interrupts from, say, an ADC, creates additional complexity for the
programmer when reasoning about shared memory, especially with respect to
idempotence.\\
    Specific to the MSP430 architecture (but common in other systems), there is
also the issue of stack integrity. On the MSP430, when an interrupt fires,
the program counter and status register are pushed onto the stack for
the handler to be able to return back to the state of the machine when the
interrupt fired. In our case, however, where power failure is the common
case, an application could experience failure while inside an interrupt
handler. This raises two issues. First, upon reboot from a power failure,
the stack will be cleared (since the stack is stored in volatile memory) and
execution will return to the handler, if we assume that the interrupt handler
is a \chain task. Upon leaving the interrupt, via the MSP430 RETI instruction),
the program's behavior will be undefined. Second, if the programmer does not
use a \chain task for the interrupt handler, it is possible that the interrupt
could fire when the system has a low amount of energy remaining, thus
rendering the system unable to finish executing the interrupt handler,
leading to a host of potential problems.

\subsubsection{Interface}
We provide the functions and macros in figure 3 as an interface to
the programmer. The standard workflow is as follows: the programmer
designates the task that should be run on an interrupt firing
with \texttt{INTERRUPT\_TASK(val, func)} where \texttt{val} is the task
index and \texttt{func} is the function that will be called upon an
interrupt firing - \texttt{val} and \texttt{func} follow the same
interface as \chain tasks. Since we assume that the handler is not
reentrant, the programmer replaces calls to \texttt{\_\_enable\_interrupt()}
with our function \texttt{enable\_interrupts()}.
The programmer's code remains unmodified, with the exception of a call to
\texttt{int\_setup\_complete()} when the \chain program has finished whatever
register initialization is necessary to setup the interrupts for the
application.\\

\begin{figure*}
\begin{minipage}[b]{0.49\textwidth}
\begin{lstlisting}
    /** Enable interrupts if int_setup_complete() has been called
      * and execution is not inside an interrupt handler */
    void enable_interrupts();

    /** To be called when setup for the interrupt handler is complete,
      * also enables interrupts */
    void int_setup_complete();

    /** Return from the interrupt handler */
    void return_from_interrupt();

    /** Designate a task to run when an interrupt fires */
    INTERRUPT_TASK(val, func)
\end{lstlisting}
\caption{Interrupt Handling Interface}\label{label-a}
\end{minipage}\hfill
\end{figure*}

\subsubsection{Design}
Working within the constraints of \chain and our multi-programming extensions,
we work to provide a layer of abstraction to the programmer through the standard
layer of tasks and channels. Notably, this includes interrupt handlers being
tasks, rather than standard functions. The alternative has been explored
by keeping specific capacitors for interrupt handlers\cite{Aware}, but this
approach fails if interrupt handlers take longer than the capacitor has
energy for. This bug handlers not completing due to power failure is
unintuitive for programmers coming from consistently powered devices and
makes certain devices unusable under intermittent power.\\

With interrupt handlers as tasks, if a power failure occurs during a
handler, execution will resume at the handler. This isn't ideal, the
standard paradigm that interrupt handlers are generally short and fast
so that interrupts won't be missed (for instance, a timer interrupt
being missed because processing the interrupt took longer than the
timer period), but we argue that it is more unreasonable to say
that interrupt handlers that take too long will not complete than
to violate the standard paradigm, since the former arbitrarily restricts
the application domain.\\

For the sake of simplicity, we assume that interrupt handlers are not
reentrant. Since reentrant handlers can be used in a non-reentrant fashion
without consequences, this assumption does not restrict the application domain.
Moreover, the benefit to reentrant handlers, the application will not miss an
interrupt firing because processing the previous one took too long, is
secondary on intermittent architectures - there is likely a higher chance that
a power failure will occur than an interrupt handler causing an interrupt to be
missed, unless interrupts are firing on an extremely fine granularity.\\

When an interrupt handler has completed, the programmer calls
\texttt{return\_from\_interrupt()}, which returns execution to the start
of the task that was running when the interrupt occurred. This contradicts
expected behavior - on continuously powered systems, execution generally
returns to the instruction after the instruction that the system received an
interrupt while processing. The continuous system model is intuitive, but
breaks under \chain's intermittence. Task idempotence in \chain only works
when execution returns to the top of a
task upon power failure. Program state (registers and stack memory) are stored
in volatile memory, if an interrupt handler returned to the middle of a task,
it would be impossible to reason about program correctness, since none of the
program state had been stored. The programmer could store program state on the
stack at the start of interrupt handlers, but this would be lost on power
failure. She could also store this program state in non-volatile memory -
trivial for the register set, but difficult and potentially extremely costly to
store the stack, as execution could be several functions deep. Even determining
how much stack to save could be prohibitively costly. Another alternative would
be to store program state in non-volatile, but this has been shown to be
prohibitively costly in most applications \cite{Aware}. \\

Interrupts - NEIL\\

\section{Implementation}
We only had to break \chain a little. We promise!
Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
Mutexes (Simple, meant for pseudo channels) - Emily\\
Interrupts (mucked with IE flags) - Neil\\
% Interrupt prologue
% Interrupt setup

\section{Results} - Emily
No really, there will be things here!
Applications/measurements\\
Blinker (x2)\\
RSA+cukoo\\
No verification\\
Memory overhead (read off compilation)\\
Performance (\# reboots)\\

\section{Future Work}
Future work is focused on saving memory by using channel and code reuse optimizations when forking new chain programs from current tasks. 
When forking a new task, a channel may be duplicated, which linearly increases the amount of memory neceesary for channels with the number of currently running programs.
It is possible to reuse channels between tasks if the number of iterations of that task are known at compile time and it can be estimated how 
many are to be running simulateously. A register allocation graph coloring scheme can then be repurposed for reusing channels allocated at compile time.
In either case the current channel naming scheme must be changed to support channel duplication, ether real or simulated.
Currently channels are allocated and named at compile time by combining the names of the input and output channel tasks. 
However, if tasks can be duplicated this naming scheme must be altered to support multiple channels of the same name and potentially multiple channels of the same name.
This can be done with a vector clock timestamp or a simple counter stored in nonvolatile memory to create unique channel names.

Additionally, it is possible to save memory by sharing arrays between tasks without the necessity of copying the array in its entirety in each channel.
To do this the concept of interrmittent tasks mutexes for shared resources and data structures would be introduced.
The structures are protected by a lock that contains the thread ID of the task that currently holds the lock.
In order to allow all tasks access to a data structure without breaking all chain semantics, a modified version of a module channel would be used in order 
for multiple tasks to potentially modify one channel. 
However, even with that change it is challenging to define sharing between threads in an intermittent context and to maintain chain semantics while implementing 
shared data structures while still maintaining correctness. 
Because channel sizes are constrained at compile time currently, it is also necessary to know at compile time all data structures, shared or private, 
that are going to be used during execution of a chain program or programs. 
It is a challenge for future work in this field to constrain memory for channels at compile time and for the compiler to know with certainty how much memory 
is necessary to allocate. Depending on the workload, the amount of forked threads would need to be constrained such that threads don't overflow nonvolatile memory at runtime.

Future work also includes optimizations to reduce overhead, such as converting the free indices array to a free indices bitmap in order to reduce the amount of 
baseline memory that the scheduler task uses. 
Re-assignation and re-use of channels at runtime would increaes overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.

\section{Conclusion}
Multithreading in intermittence {\em is} a good idea!

\balance
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography
\end{document}